Non-Vertical Cultural Transmission, Assortment, \\and the Evolution of Cooperation

### 2020-09-09 ms/main.tex
update spatial figure; formatting

We constructed an additional model in which partner association emerges from the population structure.
We model a non-overlapping population colonizing a 2D grid of size 100x100, where each site is inhabited by one individual, similarly to \citet{lewin2020rockpaperscissors}.
Each individual is characterized by its phenotype which is either a cooperator, $A$, or a defector $B$.
In each generation individuals interact with their neighbors, i.e. individuals that inhabit adjacent sites, in a prisoner’s dilemma game (\autoref{table:prisoner_payoff}) and in horizontal cultural transmission.
As before, $T_A$ and $T_B$ are the probabilities to transmit the phenotypes $A$ and $B$ during horizontal transmission.
At each generation, each individual is chosen to "initiate" interactions with probability 1/2.
Initiators then interact with a random neighbor for a prisoners’ dilemma game and a random neighbor (with replacement) for horizontal cultural transmission (as both the game and the transmission are symmetrical, the expected number of each of these interactions per individual per generation is 1).
Therefore, the effective partner association $\alpha$ in this model is the probability that the same neighbor is picked for both interactions, or $\alpha=1/m$, where $m$ is the number of neighbors. On an infinite grid, $m=8$, but on a finite grid $m$ can be lower in edge neighborhoods close to the grid border.

@@

%%% Figure: Spatial model
\begin{figure}[htb]
  \centering
    \includegraphics[scale=0.7]{spatial.pdf}
  \caption{
  \textbf{Evolution of cooperation in a spatial model.}
  The proportion of cooperators after 10,000 generations is plotted (color coded) as function of the ration between the cost and benefit of cooperation ($c/b$) on the y-axis, and the horizontal transmission rate of both phenotypes $A$ and $B$ (without transmission bias, $T=T_A=T_B$) on the x-axis. The blue curve represents the condition for the evolution of cooperation in a well-mixed population with partner selection, $c/b<\alpha T / (1-T)$(\autoref{eq:equal_transmission}), where $\alpha=1/8$.
  The difference between the results of the spatial model and the analytical condition are because in the spatial model selection is local (i.e. among neighbors, offspring are born to their parent's neighborhood) whereas in the well-mixed population model selection is global.
  Here, population size is $10,000$ (100-by-100 grid); cost of cooperation, $c=0.05$. Simulations were stopped at generation 10,000 or if one of the phenotypes fixed.
  }
  \label{fig:spatial}
\end{figure}

>>>>

We constructed an additional model in which partner association emerges from the population structure.
We model a non-overlapping population colonizing a 2D grid of size 100-by-100, where each site is inhabited by one individual, similarly to \citet{lewin2020rockpaperscissors}.
Each individual is characterized by its phenotype which is either a cooperator, $A$, or a defector $B$.
In each generation individuals interact with their neighbors, i.e. individuals that inhabit adjacent sites, in a prisoner's dilemma game (\autoref{table:prisoner_payoff}) and in horizontal cultural transmission.
As before, $T_A$ and $T_B$ are the probabilities to transmit the phenotypes $A$ and $B$ during horizontal transmission.
At each generation, each individual is chosen to "initiate" interactions with probability 1/2.
Initiators then interact with a random neighbor for a prisoners’ dilemma game and a random neighbor (with replacement) for horizontal cultural transmission (as both the game and the transmission are symmetrical, the expected number of each of these interactions per individual per generation is 1).
Therefore, the effective partner association $\alpha$ in this model is the probability that the same neighbor is picked for both interactions, or $\alpha=1/m$, where $m$ is the number of neighbors. On an infinite grid, $m=8$, but on a finite grid $m$ can be lower in edge neighborhoods close to the grid border.

@@

%%% Figure: Spatial model
\begin{figure}[htb]
  \centering
    \includegraphics[scale=0.7]{spatial.pdf}
  \caption{
  \textbf{Evolution of cooperation in a spatial model.}
  The expected frequency of cooperators in a structured population after 10,000 generations is shown (red for 0\%, green for 100\%) as function of both the ratio between the cost and benefit of cooperation ($c/b$) on the y-axis, and the horizontal transmission rate (without transmission bias, $T=T_A=T_B$) on the x-axis.
  The population evolves on a 100-by-100 grid. Selection, cooperation, and horizontal cultural transmission are all local between adjacent sites.
  The black curve represents the condition for the evolution of cooperation in a well-mixed population with partner association, $c/b<\alpha T / (1-T)$, where $\alpha=1/8$; see \autoref{eq:equal_transmission}.
  Note that in the structured population, selection is local, whereas in the unstructured population, selection is global. This can explain the small difference in the results.
  Here, population size is $10,000$ (100-by-100 grid); cost of cooperation, $c=0.05$. 50 simulations were executed per parameter set. Simulations were stopped at generation 10,000 or if one of the phenotypes fixed.
  }
  \label{fig:spatial}
\end{figure}
